\chapter{Common}
\label{chap:common}

This chapter collects definitions and results that are not specific to any one
paper: the multigraph infrastructure, and the general theory of families of
objects, Fulkerson duality, and modulus. Later papers' chapters specialize
this material (e.g.\ to spanning trees) and build their results on top of it.

\section{Multigraphs}

\begin{definition}[Multigraph]
\label{def:multigraph}
\lean{Multigraph}
\leanok
Following the paragraph preceding Section 1.1 of \feupaperref{}: throughout, $G
= (V, E)$ denotes a finite, connected multigraph with vertex set $V$ and edge
set $E$. Parallel edges are allowed; self-loops are not.

To formalize this, a multigraph on vertex type $V$ and edge type $E$ consists of
a map $\mathrm{endpoints} \colon E \to \operatorname{Sym}^2(V)$ assigning to
each edge its (unordered) pair of endpoints. By making the edges a separate type
(rather than a relation on $V$ or a subset of unordered pairs of vertices), we
implicilty allow parallel edges. They are  simply distinct elements of $E$ with
the same image under $\mathrm{endpoints}$. Finiteness, connectedness, and the
absence of self-loops are not built into the type itself. Instead, they are
added explicitly wherever a paper's results need them, matching the
``throughout this paper'' statement.

Throughout, we write $G = (V, E, \sigma)$ for a multigraph $G = (V, E)$ together
with an edge-weight function $\sigma \in \mathbb R_{>0}^E$. However, $\sigma$ is
not part of the multigraph structure itself. We use the triple notation for
modulus results, which is where the edge weights are relevant. . For the
weight-agnostic combinatorial defintions and theorems (families of objects,
densities, admissibility, the Fulkerson dual) we write the plain pair $G = (V,
E)$ instead, since none of those notions depend on $\sigma$.
\end{definition}

\begin{definition}[Spanning tree]
\label{def:spanning-tree}
\uses{def:multigraph}
\lean{Multigraph.IsSpanningTree}
\leanok
A set of edges $T \subseteq E$ is a \emph{spanning tree} of $G$ if it touches
every vertex of $V$ and contains no cycle. In particular, this immediately rules
out parallel edges, which would form a 2-cycle.

This is formalized in \texttt{Multigraph.IsSpanningTree}: $T$ is acyclic (no
loops, no parallel edges within $T$, and the simple graph obtained by forgetting
edge identities is acyclic) and that same simple graph is connected, i.e.\
touches and joins every vertex of $V$.
\end{definition}

\section{Families of objects and Fulkerson duality}

This section follows Section 1.5 of \feupaperref{}. Families of objects generalize
spanning trees (and paths, cuts, etc.) to a common framework, in which the
notion of an admissible density and its Fulkerson dual family are defined.
Spanning trees are recovered as the special case worked out in
Section~\ref{sec:spanning-tree-modulus}.

The machinery of Fulkerson duality has been developed in the context of finite
families of objects (such as the spanning trees of a finite graph). The results
in this section are stated for finite families, but the definitions are given in
a more general context that allows for countably infinite families as well.

\begin{definition}[Family of objects]
\label{def:family-of-objects}
\uses{def:multigraph}
Let $G = (V, E)$ be a multigraph. A \emph{family of objects} on $G$ is a
nonempty countable set $\Gamma \subseteq \mathbb R_{\ge 0}^E$ ($\eta$-space). In
practice, each each $\gamma \in \Gamma$ is associated with graph-theoretic
object (a spanning tree, path, cut, etc.) identified with its \emph{usage
vector}: $\gamma(e)$ is the usage of edge $e$ by $\gamma$.
\end{definition}

\begin{definition}[Density]
\label{def:density}
\uses{def:multigraph}
A \emph{density} is a point $\rho \in \mathbb R_{\ge 0}^E$ that assigns a
nonnegative cost $\rho(e)$ to each edge $e \in E$.
\end{definition}

\begin{definition}[Length of an object]
\label{def:length}
\uses{def:family-of-objects, def:density}
Let $\rho$ be a density on $G$. The \emph{length} of an object $\gamma \in
\Gamma$ with respect to $\rho$ is
\[
  \ell_\rho(\gamma) := \sum_{e \in E} \gamma(e) \rho(e).
\]
\end{definition}

\begin{definition}[Admissible density]
\label{def:admissible-density}
\uses{def:length}
A density $\rho \in \mathbb R_{\ge 0}^E$ is \emph{admissible} for a family
$\Gamma$ if $\ell_\rho(\gamma) \ge 1$ for every $\gamma \in \Gamma$. The
\emph{admissible set} of $\Gamma$ is defined as
\[
  \Adm(\Gamma) := \{\rho \in \mathbb R_{\ge 0}^E : \ell_\rho(\gamma) \ge 1 \ \forall \gamma \in \Gamma\}.
\]
\end{definition}

\begin{definition}[Fulkerson dual family]
\label{def:fulkerson-dual}
\uses{def:admissible-density}
Let $\Gamma$ be a finite family of objects on $G = (V, E)$.
The \emph{Fulkerson dual family} of $\Gamma$ is $\widehat\Gamma :=
\operatorname{ext}(\Adm(\Gamma)) \subset \mathbb R_{\ge 0}^E$, the set of
extreme points of $\Adm(\Gamma)$.
\end{definition}

\begin{theorem}[Fulkerson duality]
\label{thm:fulkerson-duality}
\uses{def:fulkerson-dual}
For any finite family of objects $\Gamma$, $\widehat{\widehat\Gamma} \subseteq \Gamma$.
\end{theorem}

\section{Modulus of families of objects}

This section follows Section 1.6 of \feupaperref{}.

\begin{definition}[Energy of a density]
\label{def:energy}
\uses{def:density}
Let $1 \le p \le \infty$ and let $G = (V, E, \sigma)$ be a multigraph. The \emph{energy} of a density $\rho$ is
\[
  \Energy_{p,\sigma}(\rho) := 
  \begin{cases}
    \sum_{e \in E} \sigma(e) \rho(e)^p & 1 \le p < \infty, \\
    \max_{e \in E} \sigma(e) \rho(e) & p = \infty.
  \end{cases}
\]
\end{definition}

\begin{definition}[$p$-modulus]
\label{def:modulus}
\uses{def:admissible-density, def:energy}
Let $G = (V, E, \sigma)$ be a multigraph.
Fix $1 \le p \le \infty$. The \emph{$p$-modulus} of a family of objects $\Gamma$
on $G$ is
\[
  \Mod_{p,\sigma}(\Gamma) := \inf_{\rho \in \Adm(\Gamma)} \Energy_{p,\sigma}(\rho).
\]
A density achieving this infimum is called \emph{extremal} (or optimal).
\end{definition}

\begin{definition}[Extremal density]
\label{def:extremal-density}
\uses{def:modulus}
A density $\rho^* \in \Adm(\Gamma)$ is \emph{extremal} (or optimal) for $\Mod_{p,\sigma}(\Gamma)$ if
\[
  \Mod_{p,\sigma}(\Gamma) = \Energy_{p,\sigma}(\rho^*).
\]
\end{definition}

\begin{theorem}[Existence and uniqueness of the extremal density]
\label{thm:modulus-extremal-unique}
\uses{def:extremal-density}
Let $\Gamma$ be a finite family of objects on $G = (V, E, \sigma)$ with
$\Adm(\Gamma) \ne \emptyset$, and fix $1 < p < \infty$. Then there is a unique
density $\rho^* \in \Adm(\Gamma)$ achieving $\Mod_{p,\sigma}(\Gamma)$.
\end{theorem}

\begin{theorem}[Fulkerson duality for modulus]
\label{thm:modulus-fulkerson-duality}
\uses{def:modulus, thm:modulus-extremal-unique}
Let $G = (V, E, \sigma)$ be a multigraph. Let $\Gamma$ be a finite family of
objects on $G$. Let $1 < p <
\infty$ and let $q$ satisfy $pq = p + q$. Then
\[
  \Mod_{p,\sigma}(\Gamma)^{1/p} \, \Mod_{q,\sigma^{1-q}}(\widehat\Gamma)^{1/q} = 1,
\]
and the extremal density $\eta^*$ for $\Mod_{q,\sigma^{1-q}}(\widehat\Gamma)$ is
related to the extremal density $\rho^*$ for $\Mod_{p,\sigma}(\Gamma)$ by
\[
  \eta^*(e) = \frac{\sigma(e) \rho^*(e)^{p-1}}{\Mod_{p,\sigma}(\Gamma)} \qquad \forall e \in E.
\]
In particular, when $p = q = 2$,
\[
  \Mod_{2,\sigma}(\Gamma) \, \Mod_{2,\sigma^{-1}}(\widehat\Gamma) = 1,
  \qquad
  \eta^*(e) = \frac{\sigma(e)}{\Mod_{2,\sigma}(\Gamma)} \rho^*(e).
\]
\end{theorem}
